---
title: "HW6"
author: "Chris Hill"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r}
#| echo: false
#| message: false

library(httr)
library(tidyverse)
library(lubridate)
library(jsonlite)

catch_error <- function(test) {
  tryCatch(
    test,
    error = function(e) {
      message("Caught error: ", e$message)
    }
  )
}
```

## Task 1 - Conceptual Questions

1. *What is the purpose of the `lapply()` function? What is the equivalent purrr function?*

    * `lapply()` is a Base R function that takes a list and applies a given function to every element of that list. It then returns a list of equal size.
    * The purrr package has an equivalent `map()` function. I am familiar with Scala, and so this is more to my liking, personally. Particularly the type checking extensions.

2. *Suppose we have a list called `my_list`. Each element of the list is a numeric data frame (all columns are numeric). We want to use `lapply()` to run the code `cor(numeric_matrix, method = "kendall")` on each element of the list. Write code to do this below! (I’m really trying to ask you how you specify `method = "kendall"` when calling `lapply()`)*

    * The arguments to `cor()` are not fed into `cor()` itself, but are instead provided as additional arguments to `lappy()`
    
```{r}
#| eval: false
result <- lapply(my_list, cor, method = "kendall")
```

3. *What are two advantages of using purrr functions instead of the Base R `apply` family?*

    * The first argument to all map functions is always the data. Thus it is "pipable"
    * `map_*()` allows us to specify the type of output. This provides us with type checking!

4. *What is a side-effect function?*

    *  Side effect effect function doesn't actually try to change the data. It just tries to produce something or alter the environment.

5. *Why can you name a variable `sd` in a function and not cause any issues with the `sd()` function?*

    * Lexical Scoping
    * Functions actually have temporary function environments.

## Task 2 - Writing R Functions

### 1. Root Mean Square Error (RMSE)

```{r}
getRMSE <- function(actual, predicted, ...) {
  sqrt(mean((actual - predicted)^2, ...))
}
```

### 2. Test getRMSE()

Simulated Test Data

```{r}
set.seed(10)
n <- 100
x <- runif(n)
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

```{r}
getRMSE(resp, pred)
```

Inject an NA into the Test Data

```{r}
resp[c(5, 10)] <- NA_real_
```

```{r}
getRMSE(resp, pred)
```

Use additional arguments to remove NA

```{r}
getRMSE(resp, pred, na.rm = TRUE)
```

### 3. Mean Absolute Error (MAE)

```{r}
getMAE <- function(actual, predicted, ...) {
  mean(abs(actual - predicted), ...)
}
```

### 4. Test getMAE

```{r}
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

```{r}
getMAE(resp, pred)
```

Inject an NA into the Test Data

```{r}
resp[c(5, 10)] <- NA_real_
```

```{r}
getMAE(resp, pred)
```

Use additional arguments to remove NA

```{r}
getMAE(resp, pred, na.rm = TRUE)
```

### 5. Wrapper for RMSE and MAE

```{r}
getMetrics <- function(actual, predicted, metrics = c("RMSE", "MAE"), ...) {

  # Validate inputs
  if (!is.atomic(actual) || !is.atomic(predicted)) {
    stop("Both inputs must be atomic vectors.")
  }
  
  if (!is.numeric(actual) || !is.numeric(predicted)) {
    stop("Both inputs must be numeric vectors.")
  }
  
  results <- list()
  
  if ("RMSE" %in% metrics) {
    results$RMSE <- getRMSE(actual, predicted, ...)
  }
  
  if ("MAE" %in% metrics) {
    results$MAE <- getMAE(actual, predicted, ...)
  }
  
  return(results)
}
```

### 6. Test getMetrics()

```{r}
resp <- 3 + 10 * x + rnorm(n)
pred <- predict(lm(resp ~ x), data.frame(x))
```

Test using this data. Call it once asking for each metric individually and once specifying both metrics

```{r}
getMetrics(resp, pred)
```

```{r}
getMetrics(resp, pred, metrics = "RMSE")
getMetrics(resp, pred, metrics = "MAE")
```

Inject an NA into the Test Data

```{r}
resp[c(5, 10)] <- NA_real_
```

Test using this data. Call it once asking for each metric individually and once specifying both metrics

```{r}
getMetrics(resp, pred)
```

```{r}
getMetrics(resp, pred, metrics = "RMSE")
getMetrics(resp, pred, metrics = "MAE")
```

Test function by passing it incorrect data

```{r}
getMetrics(data.frame(resp), pred) |> catch_error()
```

## Task 3 - Querying an API and a Tidy-Style Function

### 1. Query API with GET()

```{r}
api_key <- "d9d2b8e6433e4e2d8364a4dfa5ea23cf"
base_url <- "https://newsapi.org/v2/everything"

params <- list(
  q = "technology",    
  from = "2024-10-01",    
  apiKey = api_key
)

response <- GET(base_url, query = params)
response
```

### 2. Parse the response

```{r}
response_content <- fromJSON(rawToChar(response$content), flatten = TRUE)

articles <- pluck(response_content, "articles")

articles_df <- as_tibble(articles)
articles_df
```

### 3. Reusable Function

```{r}
queryNewsAPI <- function(subject, from_date, api_key) {
  
  base_url <- "https://newsapi.org/v2/everything"
  
  # Can only query 30 days past
  input_date <- as_date(from_date)
  max_from_date <- Sys.Date() - 30
  
  if (input_date < max_from_date) {
    warning("Setting the from_date to the 30 day max")
    from_date <- as.character(max_from_date)
  }
  
  params <- list(
    q = subject,
    from = from_date,
    apiKey = api_key
  )
  
  response <- GET(base_url, query = params)
  
  if (http_status(response)$category != "Success") {
    stop("API Request Failed")
  }
  
  response_content <- fromJSON(rawToChar(response$content), flatten = TRUE)
  
  articles <- pluck(response_content, "articles")
  
  as_tibble(articles)
}
```

```{r}
tech <- queryNewsAPI("technology", "2024-10-01", api_key)
tech
```

```{r}
health <- queryNewsAPI("health", "2024-10-01", api_key)
health
```

### 4. Summarize the source

```{r}
source_summary <- table(tech$source.name)

source_summary
```

### 5. publishedAt column

```{r}
getTimeBetweenArticles <- function(article) {
  article |>
    mutate(publishedAt = ymd_hms(publishedAt)) |>  
    arrange(publishedAt) |>  
    mutate(pub_diff = publishedAt - lag(publishedAt)) 
}
```

```{r}
my_tech <- tech |> getTimeBetweenArticles()

my_tech |> select(publishedAt, pub_diff, everything())
```

```{r}
my_health <- health |> getTimeBetweenArticles()

my_health |> select(publishedAt, pub_diff, everything())
```

### 6. Subset and Statistics

```{r}
getPublishedStats <- function(article) {
  subset <- article |>
    select(publishedAt, pub_diff)
  
  summary_stats <- map(subset, \(x) {
    c(mean = mean(x, na.rm = TRUE),
      sd = sd(x, na.rm = TRUE),
      median = median(x, na.rm = TRUE))
  })
  
  summary_stats
}
```

```{r}
my_tech |> getPublishedStats()
```

```{r}
my_health |> getPublishedStats()
```


